A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950

### 2022-02-11 paper4.Rmd
add birth month, sex and temp

---
title: "A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950"
author: "Göran Broström and Tommy Bengtsson"
date: "`r Sys.time()`"
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  bookdown::word_document2:
    toc: no
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
citation_package: natbib
classoption: titlepage
bibliography: bio.bib
titlepage: yes
biblio-style: apalike
documentclass: article
fontsize: 11pt
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{graphicx}
- \usepackage{a4wide}
#- \usepackage[tablesonly]{endfloat}
#- \renewcommand{\efloatseparator}{\mbox{}}
abstract: "The effect of extreme temperatures on infant mortality in the Umeå and Skellefteå regions 1895-1950 is studied in a biometric analysis setting. More precisely, the effect of climate and weather, measured by temperature, average and extremes, on infant mortality is investigated. It turns out that climate (average) is more important than weather (extremes), low average temperatures are more important than temporary dips in temperature, but effects are different in neonatal and postneonatal settings."
---

@@

@@

@@

>>>>

---
title: "A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950"
author: "Göran Broström and Tommy Bengtsson"
date: "`r Sys.time()`"
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
  bookdown::word_document2:
    toc: no
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
citation_package: natbib
classoption: titlepage
bibliography: bio.bib
titlepage: yes
biblio-style: apalike
documentclass: article
fontsize: 11pt
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{graphicx}
- \usepackage{a4wide}
#- \usepackage[tablesonly]{endfloat}
#- \renewcommand{\efloatseparator}{\mbox{}}
abstract: "The effect of extreme temperatures on infant mortality in the Umeå and Skellefteå regions 1895-1950 is studied in a biometric analysis setting. More precisely, the effect of climate and weather, measured by temperature, average and extremes, on infant mortality is investigated. It turns out that climate (average) is more important than weather (extremes), low average temperatures are more important than temporary dips in temperature, but effects are different in neonatal and postneonatal settings."
---

@@

#### Temperature and sex

We check the interaction between temperature and sex and period.

```{r neosexper}
fit <- tpchreg(oe(event, exposure) ~ sex * (extemp + extemp.1 + emeantemp), 
               time = age, data = winter)
dr <- drop1(fit, test = "Chisq")
dr
```

Obviously no temperature and sex interactions

@@

#### By birth month

```{r bybmonth, fig.cap = "Crude postneonatal relative (to January = 1) mortality by birth month.", cache = TRUE}
pn <- age.window(infdat, c(28/365, 1))
fit <- coxreg(Surv(enter, exit, event) ~ as.factor(birthmonth), data = pn)
plot(exp(c(0, coef(fit))), type = "b", xlab = "Birth month", ylab = "Hazard ratios",
     col = "blue")
abline(h = 1)
```

@@

#### Sex and temperature

```{r postanalwpsex, results='asis', cache = TRUE}
##tpneo <- toTpch(Surv(enter, exit, event) ~ lowTemp + highTemp + extemp + emeantemp + sex + socBranch + illeg + parity
  ##              + subreg + period, data = infdat, cuts = c(0, 1))
##winter <- gtpneo[gtpneo$summer == 0, ]
##winter$season <- factor(winter$season)

fit.postwsex <- tpchreg(oe(event, exposure) ~ sex * (extemp + extemp.1 + emeantemp), 
                     data = tabpost[!tabpost$summer, ], time = age)
dr.postwsex <- drop1(fit.postwsex, test = "Chisq")
```

Interaction tests with temperature and sex, postneonatal mortality:

```{r postanalwsex, cache = FALSE}
round(dr.postwsex, 4)
#tbl(round(dr.postw, 4), caption = "Interaction tests with temperature and period, postneonatal ##mortality.", rownames = TRUE)
```

Obviously no interactions between temperature and sex.
